---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(sdtm.terminology)
```

# sdtm.terminology <img src="man/figures/logo-wo-fonts.svg" align="right" height="139" alt="" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/patterninstitute/sdtm.terminology/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/patterninstitute/sdtm.terminology/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{sdtm.terminology}` is an R data package providing CDISC SDTM terminology.

[CDISC SDTM Controlled
Terminology](https://www.cdisc.org/standards/terminology/controlled-terminology)
refers to a standardized set of terms developed by the [Clinical Data
Interchange Standards Consortium (CDISC)](https://www.cdisc.org/) specifically
for use with the [Standard Data Tabulation Model
(SDTM)](https://www.cdisc.org/standards/foundational/sdtm). Controlled
terminology provides a consistent vocabulary for coding various types of data
that are collected during clinical trials. This standardization is crucial for
ensuring that data from different studies, sponsors, and sites can be
understood, compared, and analyzed in a consistent manner.

In addition to supplying SDTM terminology data, this package includes several
utility functions designed to facilitate the conversion between codes and terms,
as well as to verify the association of terms with specific code lists.

## Installation

``` r
# install.packages("pak")
pak::pak("patterninstitute/sdtm.terminology")
```

## Version

The current controlled terminology release bundled with this package is:

```{r}
# SDTM controlled terminology version
ct_release()
```

## SDTM terms

```{r}
# SDTM terms (other than code lists)
ct()
```

## Code lists

```{r}
# SDTM code lists
ct("list")
```

## Full SDTM terminology 

```{r}
# SDTM terminology individual terms and code list terms.
ct("all")
```


## Crosswalks

Convert terms to codes and codes to terms, and determine if a term is included
in a terminology code list:

```{r}
# Convert a code list term to respective C-code and back again
clst_term_to_code("UNIT")
clst_code_to_term("C71620")
# Are "%" and "%/s" terms from the UNIT code list?
is_term(c("%", "%/s"), clst_term_to_code("UNIT"))
# Find "%" and "%/s" codes within C71620 code list.
term_to_code(c("%", "%/s"), "C71620")
# It seems "%" is also a term in the VSRESU code list but not "%/s".
is_term(c("%", "%/s"), clst_term_to_code("VSRESU"))
# Or, equivalently, query using codes instead...
is_code(c("C25613", "C163560"), clst_term_to_code("VSRESU"))
# Find which code lists contain the terms "DNA" or "RNA"
match_code_list_from_term(c("DNA", "RNA"))
# Round trip from terms to codes to parent code list codes...
match_code_list_from_code(term_to_code(c("%", "%/s"), "C71620"))
```


